Prediction Markets for Exponential Families

### 2014-01-31 WriteUp/Original/maxent-market.tex
tidied up main folder to just have submission doc

% Adjusted to add \usepackage[numbers]{natbib}
% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
% Jan 5, 2013 - David Parkes
%
% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2013}
\acmMonth{6}
\usepackage[numbers]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{parskip}
\usepackage{booktabs}

%new commands
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\bR}{\mathbf{R}}
\newcommand{\grad}{\nabla}
\newcommand{\Exp}{\mathbf{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\mM}{\ensuremath{\mathcal{M}}}
\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\htheta}{\hat{\theta}}
%---------------------------------------------------------------------------
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{con}[thm]{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{ex}[thm]{Example}
\newtheorem{qn}[thm]{Question}
\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
\newcommand{\undefn}{\bigbreak}
\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
\newcommand{\unrmk}{\bigbreak}
%\newcommand{\proof}{\noindent{\bf Proof. }}
\newcommand{\unproof}{\hfill$\Box$\bigbreak}
\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
\newcommand{\holds}[1]{\stackrel{?}{#1}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ie} {{\it i.e. }}
\newcommand{\eg} {{\it e.g. }}
\newcommand{\etal} {{\it et al. }}
\newcommand{\E}{\mathbf{E}}
\newcommand{\betavec}{\pmb{\beta}}
\newcommand{\qvec}{\mathbf{q}}
\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
\newcommand{\family}{\mathcal F} 
\newcommand{\suff}{\pmb{\phi}} % suff stats
\newcommand{\muvec}{\pmb{\mu}}

% Document starts

\begin{document}
% Page heads
\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
% Title portion
\title{Maximum Entropy Prediction Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor}}
%%\author{GANG ZHOU
%%\affil{College of William and Mary}
%%YAFENG WU
%%\affil{University of Virginia}
%%TING YAN
%%\affil{Eaton Innovation Center}
%%TIAN HE
%%\affil{University of Minnesota}
%%CHENGDU HUANG
%%\affil{Google}
%%JOHN A. STANKOVIC
%%\affil{University of Virginia}
%%TAREK F. ABDELZAHER
%%\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

\category{J.4}{Social and Behavioral Sciences}{Economics}
\category{I.2.6} {Artificial Intelligence}{Learning}

\terms{Prediction Market Design, Machine Learning Algorithms}

\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}

\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.

% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
%%This work is supported by the National Science Foundation, under
%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
%%
%%Author's addresses: G. Zhou, Computer Science Department,
%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
%%Computer Science Department, University of Virginia; T. Yan,
%%Eaton Innovation Center; T. He, Computer Science Department,
%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
\end{bottomstuff}

\maketitle

\section{Introduction}
%------------------------------------------------------------
%Outline
%------------------------------------------------------------
%- exp family review
%- maxent scoring rule
%	- generalized lmsr
%	- does modularity (a la Hanson) still hold?
%- prediction market setup
%	- MLE with risk neutrality
%- conjugate prior market
%	- also on hyperparam?
%- exp-util
%- equilibrium for repeated trades
%- budgets and adversarial markets

We have dual goals in this paper. On the one hand, we highlight a structural similarity between prediction markets and exponential families. We see this as the syntax of our prediction market mechanism. On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments. For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
\section{Notation and Definitions}
convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
\section{Maximum Likelihood Scoring Rules}

\subsection{Preliminaries}

We consider a state space $\mX$ which may be discrete or continuous, and a measure $\nu$ over sets of states. We will typically take the state space to be  $\bR$ or $\bR^d$ together with the Lebesgue measure (perhaps restricted to some subset of states), or some finite set together with the counting measure.\footnote{We will not need to make reference to the underlying $\sigma$-algebra, but to complete the specification of the measure space we take it to be the collection of Borel sets in the case of $\bR$ and $\bR^d$, and the power set in the case of a discrete set.} A \emph{probability density} is a $\nu$-measurable function $p : \mX \rightarrow \bR_+$ that integrates to 1 over $\mX$.

Let $\mP$ denote the set of all probability densities over the given measure space. We are interested in eliciting information about some unknown density $p \in \mP$ which corresponds to the beliefs of an expert. The expert may be familiar with $p$ in its entirety, or only with certain properties of $p$.  Given a set of densities $\mD \subseteq \mP$, a \emph{property}  is a real (possibly vector-valued) function $\Gamma : \mD \rightarrow \bR^k$. With a slight abuse of terminology we call elements of its image $\Theta = \Gamma(\mD)$ \emph{properties} and we say that $p$ has property $\theta$ when $\Gamma(p) = \theta$. If $\Gamma$ is one-to-one, we instead call $\Gamma$ a \emph{parametrization} and refer to the elements of its image $\Theta$ as \emph{parameters}.

To be concrete, our outcome space could for instance be $\bR$ with $\nu$ the Lebesgue measure restricted to $[0,+\infty)$. The set $\mP$ then consists of all probability densities with support contained in the non-negative reals, and $\mD \subset \mP$ might consist of the densities of exponential distributions. An example of a property over $\mP$ is the mapping from a density to its mean.\footnote{The property may not be defined for all $p \in \mP$. In this case we implicitly restrict it to the subset of $\mP$ for which it is defined whenever its domain $\mD$ is not made explicit.} When restricted to the exponential densities $\mD$, this mapping is a parametrization. In much of this work, as in this example, the purpose of the measure $\nu$ is to implicitly encode bounds on the support of possible densities. The reason we use this approach, rather than explicitly stating bounds, is that it allows for clearer connections with maximum entropy estimation later on.

Given the set of relevant densities $\mD \subseteq \mP$, which contains the expert's belief $p$, and a property $\Gamma$ over this set, the expert is asked to report $\theta = \Gamma(p)$. The expert is rewarded according to a \emph{scoring rule} $S : \Theta \times \mX \rightarrow \bR \cup \{-\infty\}$ which pays it $S(\htheta,x)$ according to how well its report $\htheta \in \Theta$ agrees with the eventual outcome $x \in \mX$. The idea is to design the scoring rule so that the expert is incentivized to report its true belief, setting $\htheta = \theta$. 
%
\begin{definition}
Given a property $\Gamma : \mD \rightarrow \bR^k$, a scoring rule $S$ is  \emph{proper} over domain $\mD$ if for each $\theta \in \Theta$ and $p \in \mD$ with property $\theta$, we have
%
\begin{equation} \label{eq-proper}
\Exp_p[S(\theta,x)] > \Exp_p[S(\htheta,x)]
\end{equation}
%
for all $\htheta \neq \theta$. If the domain is $\mD = \mP$, the set of all possible densities, then we say the scoring rule is \emph{strongly proper}.
\end{definition}
%

\begin{itemize}
\item Note that the inequality is strict, in the literature these rules are called \emph{strictly} proper, but we only consider strictly proper rules here so don't make this qualification.
\item Lambert et al.'s rules are actually strongly proper, however we consider scoring rules over vector valued properties. Lambert et al only develop these by adding together different scoring rules for each property. We go further in our constructions and develop rules that are not necessarily separable.
\item It is useful to consider rules that are not strongly proper because we might have constructions for these for properties where no strongly proper rule is known.
\item It is important to understand the informational requirements placed on the expert. Under a proper scoring rule, the expert does not need to be aware of the entire density $p$ to be incentivized to report its property $\theta$, since $\theta$ is an optimal report no matter what the density might be. The agent only needs to agree that the density indeed lies in $\mD$.
\item For instance, if we are considering densities with support on $[0,+\infty)$, and $\mD$ is the set of lognormal densities, then a proper scoring rule over $\mD$ for the mean does not require the agent to know anything beyond the mean (e.g., it does not need to know the variance).  
\item A strongly proper scoring rule goes further; with such a rule we do not require the expert to believe the density is drawn from some subset $\mD$, only that it agrees with the support implied by the base measure $\nu$. Therefore, a strongly proper scoring rule for the mean here would elicit the mean no matter what kind of density over the positive reals might apply (e.g., exponential, Pareto, lognormal).
\item In practical scenarios this seems like an important extension, because while it might be clear which states are possible, it might be difficult to ensure the expert agrees that the probability density for states falls within some restricted family $\mD$.
\end{itemize}

\subsection{Maximum Likelihood}

%
\begin{theorem} \label{maxlike}
The logarithmic scoring rule defined by
%
\[ S(\theta, x) = \log p(x;\theta) \]
%
is weakly proper for $\mF$ over $\Theta$ if and only if the maximum likelihood estimate for $\theta$ is consistent. Furthermore, if $\mM$ is an alternative parameter space such that there exists a bijection $t : \mM \rightarrow \Theta$, then $\bar{S}(\mu, x) = S(t(\mu), x)$ is weakly proper for $\mF$ over $\mM$. 
\end{theorem}
%

%
\begin{example}
Suppose $X$ is supported on $[0,+\infty)$ and follows an exponential distribution with density $f(x;\lambda) = \lambda e^{-\lambda x}$ parametrized by a rate $\lambda > 0$. The mean $\mu$ is related to the rate via the one-to-one mapping $\mu = 1/\lambda$, so the density can alternatively be parametrized by the mean. According to Theorem~\ref{maxlike} the following is a weakly proper scoring rule for the mean of an exponential distribution:
%
\begin{equation} \label{exp-scoring}
S(\mu, x) = -\frac{x}{\mu} - \log \mu.
\end{equation}
%
To verify this, suppose the agent believes $X$ follows an exponential distribution with rate $\lambda_0$, equivalently mean $\mu_0 = 1/\lambda_0$. By the one-to-one relation between rate and mean, choosing $\mu$ to maximize the expected score~(\ref{exp-scoring}) is equivalent to choosing $\lambda$ to maximize $\Exp{-\lambda x + \log \lambda} = -\lambda \mu_0 + \log \lambda$. The latter is strictly concave in $\lambda$, and using basic calculus its unique global maximum is $\lambda^* = 1/\mu_0$, which therefore leads the agent to report $\mu = \mu_0 = 1/\lambda^*$ under scoring rule~(\ref{exp-scoring}). 
\end{example}
%

%
\begin{example}
Suppose that $X$ is supported on $[1,+\infty)$ and follows a Pareto distribution with density $f(x;\alpha) = \alpha/x^{\alpha+1}$ parametrized by an index $\alpha > 0$. The mean $\mu$ is related to the index via the one-to-one mapping $\mu = \frac{\alpha}{\alpha-1}$, so the density can alternatively be parametrized by the mean. Theorem~\ref{maxlike} gives the following weakly proper scoring rule for the mean of a Pareto distribution with support on $[1,+\infty)$:
%
\begin{equation} \label{pareto-scoring}
S(\mu,x) = \log \frac{\mu}{\mu-1} - \left(\frac{\mu}{\mu-1} + 1\right) \log x.
\end{equation}
%
To verify this, suppose the agent believes $X$ follows a Pareto distribution with index $\alpha_0$, equivalently mean $\mu_0 = \frac{\alpha_0}{\alpha_0-1}$. By the one-to-one relation between the index and mean, choosing $\mu$ to maximize the expected score~(\ref{pareto-scoring}) is equivalent to choosing $\alpha$ to maximize $\Exp{\log \alpha - (\alpha+1)\log x} = \log \alpha - (\alpha+1)\Exp{\log x}$. The latter is strictly concave in $\alpha$, and using basic calculus its unique global maximum is $\alpha^* = 1/\Exp{\log X} = \alpha_0$, where the last equality follows from the fact that $\log X$ has an exponential distribution with rate $\alpha$ when $X$ is Pareto with rate $\alpha$. Therefore~(\ref{pareto-scoring}) leads the agent to report $\mu = \mu_0 = \frac{\alpha^*}{\alpha^* - 1}$. 
\end{example}
%

\begin{itemize}
\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal. 
\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics. Explain what an m-estimator looks like and how there are general consistency results for those as well.
\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters. (Just a hunch.)
\end{itemize}
\section{Maximum Entropy Market Making}

@@

Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
%
\begin{eqnarray}
\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
 & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
\end{eqnarray}
%
The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution. Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report. The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$. The program has a convex objective and linear constraints. %For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized. 
%
\begin{theorem} \label{thm:scoring}
Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
%
\begin{equation} \label{scoring}
S(\mu,x) = a_x + b\log p(x;\mu),
\end{equation}
%
where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
\end{theorem}
%
Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$

Note that the maximum entropy distribution is an exponential family distribution. 
 Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.

 We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
   Or 
\begin{equation}\label{eq:proper}
\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
\end{equation}

   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
   Rearranging the terms,
   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
   which is the first order condition on the convexity of $\psi(\cdot)$.
 \unproof

 \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters. That is,
   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
   \unrmk

\noindent
This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.  As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution. For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}. We provide here two examples in more depth.

@@

\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable. In particular, we assume that the outcome is drawn from an exponential family distribution. Each of the traders has access to a series of points drawn from this distribution. In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family. Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.

%We will now provide a more formal description of the problem statement.
%\subsection{Problem Statement}
%\begin{description}
%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.  In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR. Define a correspondence between the state of the prediction market and the output of the learning algorithm.
%\end{description}

For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$. The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.

We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution. %Given  under some interpretation of what information the participating agents have, and how they bet.

For a given exponential family distribution, the prediction market is defined as follows:
\begin{description}
\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$. This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples. The trader trades in the market to maximize her expected payoff under her believed distribution.
\item[Securities and their payoff]
For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn. We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative. 
\end{description}

We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded. If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.

\paragraph{Arbitrage-free Property}
We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions. First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.

Now, consider any trade (or sequence of trades) that moves the market from an initial state of $\qvec$ to a final state of $\qvec + \qvec'$. The cost of this trade is $C(\qvec + \qvec') - C(\qvec)$. The cost function of our market is the log-partition function of an exponential family, and thus, $C(.)$ is a strictly convex function~\cite{WainJordan08}. Thus, we have the inequality:
\[
   C(\qvec + \qvec') - C(\qvec)  > \qvec' \nabla C|_\qvec = \qvec' \mu,
 \]
 where $\mu$ is the mean sufficient statistics vector for the distribution with parameters $\qvec$.
The payoff due to this trade, with outcome $x$, is $\qvec'.\phi(x)$. 
Now, we observe that under the distribution with parameters $\qvec$, the expected payoff is $\qvec' E[\Phi(x)] = \qvec' \mu$. Thus, under this distribution over outcomes $x$, the expected profit is strictly less than $0$. This is only possible if the
realized profit is less than $0$ for at least one outcome $x$. As this statement is true for every $\qvec$ and $\qvec'$, and the profits are path-independent, the market is arbitrage-free.

%Let $\qvec^*=(q^{*}_{0},q^{*}_{1},\ldots,q^{*}_{k})$ be the number of shares of each security held by the traders. First we note that under the assumption of perfectly rational, risk-neutral traders with infinite budget and  beliefs as indicated above, the gradient of the cost function at this point is
 %$$\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}=\mu_{i}$$
%where $\mu_{i}$ is believed to be the expected payoff of the $i^{th}$ security by the last trader who traded in this market. We show below that the instantaneous price of security $i$, $\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$ at any $\qvec=\qvec^{*}$, can be written as $\left(\frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}$. This means that the instantaneous price of the $i^{th}$ security only depends on the difference between the number of held securities of $i$ and 0. This ensures that if there exists an instantaneous price that reflects the belief of a trader at any vector $\qvec^{*}$, there must exist a $\qvec'$  that also reflects his beliefs and with the desirable property that $\forall i=0,1,2,\ldots,k\ q_{i}'\geq 0$. This additional property ensures that no trader will be required to short sell securities to move the market state to  reflect his beliefs.

\paragraph{Market State and Natural Parameters}
We define the interpretation function 
\begin{eqnarray}\label{eqn:int}
I(\qvec)&=&(q_{1}-q_{0},q_{2}-q_{0},\ldots,q_{k}-q_{0})\nonumber\\
&=&(\beta_{1}, \beta_{2},\ldots, \beta_{k})=\betavec
\end{eqnarray}
This allows us to interpret the state of the market in terms of a prediction on the natural parameters of the distribution.

\begin{theorem}
The natural parameter vector corresponding to the interpretation of the market state given by Equation \ref{eqn:int} is the vector of their maximum likelihood estimates. Further, this interpretation is unique.
\end{theorem}
\begin{proof}
We observe that for $i=1,\ldots,n$,
\begin{eqnarray*}
\mu_{i}&=&\left(\frac{\partial C(\qvec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int \exp(\qvec^{T}\Phi(x))dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}\phi_{0}(x)+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}(a-\sum_{i=1}^{n}\phi_{i}(x))+\sum_{i=1}^{n}q_{i}\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log\int e^{q_{0}a} e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial\log (e^{q_{0}a} \int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial (q_{0}a)}{\partial q_{i}} \frac{\partial\log (\int e^{\sum_{i=1}^{n}(q_{i}-q_{0})\phi_{i}(x)}dx)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\frac{\partial \beta_{i}}{\partial q_{i}}\right)_{\qvec=\qvec^{*}}\\
&=&\left(\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}\right)_{\rm \betavec=\betavec^{*}}
\end{eqnarray*}
Thus, this choice of parameters satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$. We will now argue that this vector is unique.

If the exponential family is represented so that there is a unique parameter vector associated with each distribution, the representation is said to be minimal. The Bernoulli, Gaussian, and Poisson distributions all have minimal representations. Now, for an exponential family whose representation is minimal, the gradient is an injection \cite[Prop. 3.2]{WainJordan08}. Thus, there is a unique parameter vector $\betavec$ that satisfies $\frac{\partial \lpf(\betavec)}{\partial \beta_{i}}=\mu_{i}$, for each $i\in\{1,\ldots,n\}$. Thus, if the $\mu_{i}$'s correspond to the empirical means, since our choice of $\betavec$ satisfies the equality, it must also be the vector corresponding to the MLE.
\end{proof}

Thus, we have shown that this prediction market (along with its strategic traders) aggregates trader information in a way that the market state can be interpreted as a predictive distribution (the maximum likelihood distribution) over an infinite outcome space.

It is worth noting here that if we can bound the parameter space and the sufficient statistics of the distribution, the market maker defined here has bounded worst case loss: The probability density, and hence log loss, would then be bounded for all market states that could be reached and all outcomes. In fact, for many practical applications, a distribution with unbounded parameter space is an approximation for a true space which itself may be bounded. In this case, worst case loss would thus be bounded.

\subsection{The Exponential Family Market Mechanism in Adversarial Markets}
In the previous section, we saw that we can define a cost-function based prediction market so that the aggregated belief of the traders represents the maximum likelihood estimate of the natural parameters of the true exponential family distribution.

In this section, we consider the the prediction market setup with traders that may be either informative or malicious. The malicious traders may want to inject faulty information into the market. The informative traders on the other hand receive points drawn from the true distribution on which they base their beliefs.

We will show that if we are able to impose finite initial budgets on the traders and control the market prices based on these budgets, then  it possible to set up the market so that it is prohibitive for damaging traders to participate in the market. Further, the informative traders can be shown to have expected growth in budget so that they are eventually able to move the market prices without restriction.

\subsubsection{Budget-limited Aggregation}
Imposing budget limits on the traders will allow us to control the amount of influence any one trader can have on moving the market prices. We will also satisfy an additional requirement that no trader has negative budget at any point of participation in the market. This is achieved by restricting the movement of the market and hence influencing the cost incurred by the trader. Recall that the payoff in this market is non-negative and hence the only adverse influence on a trader's budget is the cost of movement of the market state.

>>>>

%%%
%%% Maximum Entropy Market Making
%%%

\section{Maximum Entropy Market Making}

@@

Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
%
\begin{eqnarray}
\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
 & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
\end{eqnarray}
%
The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution. Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report. The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$. The program has a convex objective and linear constraints. %For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized. 
%
\begin{theorem} \label{thm:scoring}
Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
%
\begin{equation} \label{scoring}
S(\mu,x) = a_x + b\log p(x;\mu),
\end{equation}
%
where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
\end{theorem}
%
\noindent
This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.  As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution. For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}. We provide here two examples in more depth.

@@

%\footnote{There is a lack of rigor here, because there might be more than one optimal Lagrange multiplier for a given $\mu$, so the function $\theta(\mu)$ may not be well-defined. it is well-defined if the representation $\phi$ is minimal. The mapping from $\theta$ to $\mu$ is well-defined and is given by $\nabla C$.}
